# Typography and Formatting {.unnumbered}

To the extent possible, I have tried to follow the following typographic and formatting conventions throughout the course modules.

- Program names are written in ***bold italics***:

    - ***R***
    - ***RStudio***

- Functions, commands, and ***R*** code are written as `inline code`, e.g., `x <- 5`, or in code blocks:

```{r eval = FALSE}
x <- 5
print(x)
```

- Code output appears in a colored cell, prefaced by "##":

```{r echo= FALSE, eval = TRUE}
x <- 5
print(x)
```

- Package names are written in {curly braces}:

    - {tidyverse}
    - {lubridate}

- Variable names (for vectors, data frames and other tabular data, etc.) are written in **bold**:

    - **x**
    - **y**
    - **airline_flights**

- Column names within a data frame are also written in **bold**, whether referred to separately or as part of the table:

    - **origin**
    - **destination**
    - **airline_flights\$origin**
    - **airline_flights\[\["destination"\]\]**

- Filenames and file types are written in "quotation marks":

    - "myData.R"
    - ".csv"

- Full file path names are written in "quotation marks":

    - "/Users/Tony/Desktop/myData.R"
    - "\~/Desktop/myData.R"
    - "C:\\Documents and Settings\\Anthony Di Fiore\\Desktop\\myData.R"

- Directory names are written as `inline code` followed by a slash:

    - `img/`, `src/`

- Important concepts are written in *italics* when first referred to:

    - *working directory*
    - *environment*
    - *namespace*

- Menu names and menu commands are written in **bold**:

    - **File \> New File \> R Script**

- Argument values or values to be replaced are written in lowercase text between \<angle brackets\>, where that entire text, brackets included, should be replaced with the text being asked for:

    - `setwd("<your working directory>")`
    - "\<your file name\>.csv"

- Constants are written in italics:

    - *pi*

- Argument names and assignments are written as inline code:

    - Set `na.rm=TRUE`
    - Set `data=df`
    - Set `filename="~/Users/Tony/Desktop/output.csv"`

- Names of ***RStudio*** panes are written in **bold**:

    - **Console**
    - **Environment/History**

- Names of ***RStudio*** tabs within panes are written in *italics*:

    - *History*
    - *Plots*
    - *Help*
    - *Git*

- Names of dialog boxes are written in *italics*:

    - *Global Options*

- Button names and sections within dialog boxes are written in quotation marks:

    - "OK"
    - "Cancel"

- Check box names within dialog boxes are written in `inline code`:

    - `Restore .RData into workspace at startup`

- ***R*** object class names are written in **bold**:

    - **numeric**
    - **function**

- ***git*** repository branch names are written as `inline code`:

    - `main`
    - `origin/main`

- Full URLs/links are written in `inline code` and may include hyperlinks:

    - [`https://cran.r-project.org/`](https://cran.r-project.org/)
    - [`https://difiore.github.io/ada-2024/`](https://difiore.github.io/ada-2024/)
