---
title: "Notes on Plotting"
author: "Anthony Di Fiore"
date: last-modified
format:
  html:
    self-contained: true
    embed-resources: true
    css: css/style.css
    theme:
      - cosmo
    code-fold: true
    code-tools: true
    toc: true
    toc-depth: 3
execute:
  echo: true
  warning: false
  message: false
  error: true
knitr:
  opts_chunk:
    fig.width: 6
    fig.height: 4
    fig.retina: 2
    fig.align: center
    fig.dpi: 144
    out.width: "80%"
    cache: false
    dev: "png"
---

You cannot reliably add to a {base} R plot across code chunks in either a ***Quarto*** or ***R Markdown*** document. {base} graphics are tied to a graphics device (output window) that is opened and closed within a chunk. When the chunk ends, the device is closed and the plot is gone. Within the same code chunk, however, you can add to the same plot.

{ggplot2} graphics work a little differently. When you create a plot object and assign it to a variable, you can reference that variable across code chunks and add to it incrementally in different code chunks.

# Load {tidyverse} and data

```{r}
#| warning: false
library(tidyverse)
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/KamilarAndCooperData.csv"
d <- read_csv(f, col_names = TRUE)
```

# {base} R example

::: callout-note
Plotting all in one code chunk...
:::

```{r}
boxplot(data = d, log(Body_mass_female_mean) ~ Family)
stripchart(data = d, log(d$Body_mass_female_mean) ~ Family,
             method = "jitter",
             col = "blue",
             vertical = TRUE,
             add = TRUE)
```

# {ggplot2} example

::: callout-note
Plotting split across multiple code chunks
:::

::: callout-tip
# Tip
create plot object
:::

```{r}
p <- ggplot(
   data = d,
   aes(x = Family, y = log(Body_mass_female_mean))
   )
```

::: callout-tip
# Tip
add boxplot
:::

```{r}
p <- p + geom_boxplot(na.rm = TRUE)
```

::: callout-tip
# Tip
add points
:::

```{r}
p <- p + geom_jitter(color = "blue", width = 0.1, na.rm = TRUE)
```

::: callout-tip
# Tip
change theme
:::

```{r}
p <- p + theme_bw()
```

::: callout-tip
# Tip
tweak theme by rotating X axis labels
:::

```{r}
p <- p + theme(axis.text.x = element_text(angle = 90))
```

::: callout-tip
# Tip
rename Y axis
:::

```{r}
p <- p + ylab("log(Female Body Mass)")
```

::: callout-tip
# Tip
actually visualize plot
:::

```{r}
p
```


